---
title: "Project"
author: "Joaquin Sanchez"
subtitle: "Investigation into Chicago taxis"
echo: false
bibliography: references.bib
---

## Packages

```{r}
#| warning: false
#| message: false
library(tidyverse)
library(tidymodels)
library(scales)
library(modeldata)
library(ggbeeswarm)
```

## Question 1

```{r}
taxi
```

```{r}
taxi |> 
  filter(is.na(distance))
```

```{r}
#| label: histogram-distance
ggplot(taxi, aes(x = distance, fill = tip)) +
  geom_histogram(binwidth = 250)
```

```{r}
#| label: boxplot-distance
ggplot(taxi, aes(x = distance, y = tip, fill = tip)) +
  geom_boxplot(show.legend = FALSE)
```

```{r}
#| label: density-distance
ggplot(taxi, aes(x = distance, fill = tip, color = tip)) +
  geom_density(alpha = 0.5)
```

```{r}
#| label: violin-distance
ggplot(taxi, aes(x = distance, y = tip, color = tip, fill = tip)) +
  geom_violin(alpha = 0.2, show.legend = FALSE)
```

```{r}
#| label: jitter-distance
set.seed(1234)
ggplot(taxi, aes(x = tip, y = distance)) +
  geom_jitter(alpha = 0.5)
```

The graphical analysis reveals a clear relationship between trip distance and tipping behavior. While most passengers leave a tip regardless of journey length, the proportion of non-tippers declines markedly as trip distance increases, and long-distance trips consistently exhibit high tipping rates. Consequently, we conclude that longer taxi journeys are associated with a lower likelihood of passengers foregoing a tip.

## ndefinedQuestion 2

```{r}

```
