---
title: "Project"
author: "Joaquin Sanchez"
subtitle: "Investigation into Chicago taxis"
echo: false
bibliography: references.bib
---

## Packages

```{r}
#| warning: false
#| message: false
library(tidyverse)
library(tidymodels)
library(scales)
library(modeldata)
library(ggbeeswarm)
```

## Question 1

```{r}
taxi
```

```{r}
taxi |> 
  filter(is.na(distance))
```

```{r}
#| label: boxplot-distance
ggplot(taxi, aes(x = distance, y = tip, fill = tip)) +
  geom_boxplot(show.legend = FALSE) +
  theme_minimal() +
    labs(
    title = "Boxplot tip vs distance covered",
    subtitle = "Taxi service in Chicago",
    y = "Tip",
    x = "Distance in odoter miles"
  )
```

```{r}
#| label: density-distance
ggplot(taxi, aes(x = distance, fill = tip, color = tip)) +
  geom_density(alpha = 0.5) +
  theme_minimal() +
    labs(
    title = "Density plot tip vs distance covered",
    subtitle = "Taxi service in Chicago",
    y = "Density",
    x = "Distance in odoter miles"
  )
```

```{r}
#| label: violin-distance
ggplot(taxi, aes(x = distance, y = tip, color = tip, fill = tip)) +
  geom_violin(alpha = 0.2, show.legend = FALSE) +
    theme_minimal() +
    labs(
    title = "Violin plot tip vs distance covered",
    subtitle = "Taxi service in Chicago",
    y = "Tip",
    x = "Distance in odoter miles"
  )
```

```{r}
#| label: jitter-distance
set.seed(1234)
ggplot(taxi, aes(x = tip, y = distance)) +
  geom_jitter(alpha = 0.5) +
      theme_minimal() +
    labs(
    title = "Jitter plot tip vs distance covered",
    subtitle = "Taxi service in Chicago",
    y = "Distance in odoter miles",
    x = "Tip"
  )
```

The graphical analysis reveals a clear relationship between trip distance and tipping behavior. While most passengers leave a tip regardless of journey length, the proportion of non-tippers declines markedly as trip distance increases, and long-distance trips consistently exhibit high tipping rates. Consequently, we conclude that longer taxi journeys are associated with a lower likelihood of passengers foregoing a tip.

## Question 2

```{r}
#| label: boxplot-company
taxi_copy <- taxi |> 
  mutate(tip_encode = if_else(tip == "yes", 1L, 0L))

Tips_by_Company <- taxi_copy |> 
  group_by(company) |> 
    summarize(total_tips = sum(tip_encode, na.rm = TRUE))

Tips_by_Company
```

```{r}
#| label: histogram-tips
ggplot(Tips_by_Company, aes(x = total_tips, y = company)) +
  geom_bar(stat = "identity") +
    theme_minimal() +
    labs(
    title = "Total count of tips for each company",
    subtitle = "Taxi service in Chicago",
    y = "Company name",
    x = "Total count of tips"
  )
```

There is no evidence to support that passengers tend to tip more for the company Chicago Independents than the others companies.
